### 2020-04-27 paper/ms.tex
Update ms.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt,twocolumn,tighten]{aastex62}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

\newcommand{\ptfo}{PTFO$\,$8-8695}
\newcommand{\ptfob}{PTFO$\,$8-8695b}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{\today}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{Seeing Double: PTFO$\,$8-8695}

%\objectname{PTFO 8-8695}
%\objectname{PTFO 8-8695b}
%\objectname{CVSO 30}

\begin{document}

\defcitealias{bouma_wasp4b_2019}{B19}

%\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
% \title{Seeing Double: Still Against the Planetary Interpretation of
% PTFO$\,$8-8695b}
% \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
% Planet}
%\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
%PTFO$\,$8-8695b is Not a Planet}
%\title{Double Trouble: Transient Dips and Photometric Binarity in
%PTFO$\,$8-8695}
\title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
PTFO 8-8695}
\title{PTFO\,8-8695: A binary star, not a hot Jupiter host}

\correspondingauthor{L. G. Bouma}
\email{luke@astro.princeton.edu}

%
% key authors:
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J.\,N.\,Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\begin{abstract}
  PTFO$\,$8-8695 (CVSO\,30) is a star in the 7--10 million year
  old Orion-OB1a cluster that shows brightness dips that resemble planetary transits.
  While strong evidence against the planet hypothesis has already been provided,
  the possibility remains debated in the literature.
  To obtain further clues, we inspected data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS) and the ESA Gaia mission.
  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group, and independently suggest that it is an
  astrometric binary.  The TESS lightcurve shows two different photometric
  periods. The variability is dominated by a sinusoidal signal with a period
  of 11.96\,$hr, presumably caused by stellar rotation.  Also
  present is a 10.74-day signal consisting of a not-quite sinusoid interrupted
  by 1.2\% hour-long dips, the type of signal previously
  intepreted as planetary transits.
  The phase of the dips is nearly 180$^\circ$ away from the phase of the originally reported dips,
  which, as noted previously, makes them difficult to explain as planetary transits.
  Instead, we believe that PTFO$\,$8-8695 is a pair of young
  and rapidly rotating M dwarfs, one of which shows the same ``transient-dipper''
  behavior that been seen in at least 5 other cases.
  The origin of these transient dips is still unknown but likely involves circumstellar material.
\end{abstract}

\keywords{
	Exoplanet evolution (491),
  Pre-main sequence stars (1290),
	Stellar ages (1581),
	Stellar rotation (1629),
	Variable stars (1761),
  Low mass stars (2050)
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\epigraph{
	Nubble [n]: a small lump or protuberance.
}

\section{Introduction}

We wish \ptfob\ were a planet. It would be quite exceptional. It would
be the youngest known hot Jupiter
\citep{van_eyken_ptf_2012}, orbiting a T Tauri star in the Orion-OB1a cluster.
It would have the shortest orbital period (10.7 hours) of any hot Jupiter.
With such a short period, it would
probably be filling its Roche lobe, and actively
losing mass to its host star.  Not only that, but the rapidly-rotating
host star is probably oblate enough to torque the planet's orbit
into and out of the transiting configuration on a time scale of years
\citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.

Another first would be the direct detection of H$\alpha$ emission from
the planet itself \citep{johnskrull_h_2016}.  In addition to the
chromospheric H$\alpha$ emisson, it seems that there is an additional
H$\alpha$ emission with radial velocity variations in phase with the planetary
orbit.  The average velocity width of the excess H$\alpha$ emission is
87$\,$km$\,$s$^{-1}$, and its equivalent width is 70-80\% that of the
stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
explanation is that the emission is from hot material flowing away
from the planet \citep{johnskrull_h_2016}.

However, the observed signals have some peculiarities that
make the planet seem even more unusual, to the point that they cast into doubt the premise
that \ptfob\ is real.
First, the transit-like brightness dips are about
three times deeper in optical bandpasses ({\it e.g.,} $g$-band) than
in the near-infrared ({\it e.g.}, $z$-band)
\citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.  An ordinary
atmosphere expected for a Jovian planet would not lead to such a strong
color-dependence of the transits.  
Second, the planet does not seem to emit as much infrared radiation
as would be expected for such a hot Jovian planet \citep{yu_tests_2015}.
Third, despite measurement attempts by multiple investigators, \ptfob\ does
not seem to show the Rossiter effect at the amplitude expected given
the rapid stellar rotation and large planet size
\citep{yu_tests_2015,ciardi_followup_2015}.
Fourth, the phase of the dips within the overall period of photometric variability
has changed drastically over the years since their initial discovery.
To counter these objections, it has been proposed 
that the planet may be much smaller than Jupiter and that the dips
are produced by dust clouds emitted from the planet \citep{tanimoto_evidence_2020}.

A separate issue is that the brightness dips
change shape over many orbital cycles. This was initially explained by
\cite{barnes_measurement_2013} as the natural effects of gravity darkening
However, \cite{howarth_reappraisal_2016} argued
that the necessary amplitude of gravity darkening is too large to be realistic, given the
spectroscopically-determined rotation velocity.
Additionally, as the
gravity-darkened star precessed about its rotation axis, it would 
show photometric variability that has not been observed.

While the planetary interpetation clearly faces challenges,
there is no completely satisfactory alternate explanation.  High-latitude accretion hotspots
might produce the observed H$\alpha$ variability, but require
fine-tuning to produce dips of the appropriate duration. Furthermore,
\ptfo\ does not have an infrared (IR) excess associated with the
presence of dust at the inner edge of a primordial disk \citep[{\it
e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
cold, struggle to produce photometric features as short as some
of the observed dips.

A relevant fact is that between 0.1\% and 1\% of rapidly
rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
show short-duration dips as part of their overall periodic variability
\citep{rebull_usco_2018}.  The dips can persist over months, but their
depths often vary, and sometimes change immediately after stellar
flares.  The explanation proposed by \citet{stauffer_orbiting_2017}
and \citet{david_transient_2017} to explain this novel class of
variable stars is that a circumstellar cloud of gas is orbiting
near the co-rotation radius.  To this point, though, it has not
been clear if this explanation applies to
\ptfo, because the determination of
the stellar rotation period has been somewhat ambiguous
\citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two different periodic signals, which we analyze in
Section~\ref{sec:tess}.
The Gaia data, analyzed in Section~\ref{sec:gaia}, show that
\ptfo\ is a photometric binary, and suggest it also an astrometric binary.
We discuss the pieces of the puzzle
in Section~\ref{sec:discussion}, and summarize the situation in
Section~\ref{sec:conclusions}.
% JNW: this kind of stuff is for the abstract/conclusion, not intro.
%The phase of the dip within the seems
%to have instantaneously changed over the past decade.  In addition,
%other young stars show lightcurve morphologies similar to the
%short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
%binary M dwarf in which one star shows a rotation signal, and the
%other shows transient dipping potentially caused by eclipses of
%lasma at the Keplerian co-rotation radius.

\section{The Data}
\label{sec:observations}

% JNW: How about more descriptive labels? f --> Total flux
% fl = Longer-period signal
% fs = Shorter-period signal
% last panel, Residuals

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 19).}
    {\it Top}: The ``raw'' (\texttt{PDCSAP} mean-subtracted) relative flux
    data. The beat period of 4.48 days is visible by eye.  The
    blue curve is a model including 2 harmonics
    at the longer period $P_{\rm \ell}$, plus 2 harmonics and a transit
    at the shorter period $P_{\rm s}$.
		{\it Upper middle}: Longer-period signal, equal to the raw signal
		minus the shorter-period signal.
		{\it Lower middle}: Shorter-period signal, equal to the raw signal
		minus the longer-period signal.
		{\it Bottom}: residual.  The data are binned from 2 to 10 minute
		cadence as a convenience for plotting and fitting.
		\label{fig:splitsignal}
	}
\end{figure*}

\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 20).}
		Same format as Figure~\ref{fig:splitsignal}.
		\label{fig:splitsignalii}
	}
\end{figure*}

\subsection{TESS Observations}

\ptfo\ [also known as CVSO\,30; \cite{briceno_cida_2005}}]
was observed by TESS with Camera 1, CCD 1, from December 15,
2018 until January 6, 2019, during the sixth sector of science operations
\citep{ricker_transiting_2015}.  The star was designated TIC\,264461976
in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
The pixel data for an $11\times11$ array surrounding \ptfo\ were
averaged into 2-minute stacks by the onboard computer.  Each
2048$\times$2048 image from the CCD was also averaged into 30-minute
stacks, and saved as a ``full frame image'' (FFI).

The 2-minute stacks for \ptfo\ were reduced to lightcurves by the
Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  We mainly used the
Presearch Data Conditioning (PDC) lightcurve.  The PDC lightcurve
aperture used pixels chosen to maximize the SNR of the total flux of
the target \citep{smith_kepler_apertures_2017}.  Non-astrophysical
variability was removed by fitting out trends common to many stars
\citep{smith_kepler_PDC_2017}.

As an independent check on the 2-minute SPOC lightcurve, we
examined the lightcurve based upon 30-minute image stacks which was
produced as part of the Cluster
Difference Imaging Photometric Survey (CDIPS;
\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve of choice used a
circular aperture with radius 1 pixel.
% JNW: What does that mean? You used fractional pixels in calculating the pixel sums?

To clean the data, we removed all points with non-zero quality flags,
which indicate known problems
\citep[{\it e.g.},][]{tess_data_product_description_2018}.  We also
masked out the data from the first and last 6 hours of each orbit, since there are
often systematic red noise in the photometry during those times.  Both the CDIPS and PDC
lightcurves showed a clear discontinuous jump in the last few days
of orbit 20, which seemed likely to be an instrumental systematic effect.  We
correspondingly masked out the data with timestamps ranging from BJD\,2458488.3 until
the end of the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
quality-flag cut removed 854 points; masking the orbit edges removed an
additional 716 points; and removing the data from the final few days of orbit 20 removed an
additional 1079 points.  After cleaning, 83\% of the initial flux
measurements remained.

We normalized the lightcurve by dividing out the median flux, and then opted
to subtract 1.0 to set the median value to zero, which simplified
subsequent interpretation.  Many
of these and subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}.

\subsection{Gaia Observations}

\subsubsection{Astrometric measurements}

Between July 25, 2014 and May 23, 2016, Gaia measured about 300
billion centroid positions of 1{.}6 billion stars
\citep{gaia_collaboration_gaia_2016,lindegren_gaiasoln_2018,gaia_collaboration_gaia_2018}.
In the Gaia second data release (DR2), these CCD observations were
used to determine positions, proper motions, and parallaxes of the
brighest 1{.}3 billion stars, including \ptfo\
\citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good'' observations, i.e., observations that were not strongly down-weighted in
the astrometric solution.  \ptfo\ was assigned the Gaia DR2 identifier
3222255959210123904.  It photometric brightness was measured using
selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed
the pipeline parameters for \ptfo\ using the Gaia
archive\footnote{\url{gea.esac.esa.int/archive/}}.

The majority of Gaia's derived parameters for \ptfo\ agreed with
expectations from former studies
\citep{briceno_cida_2005,van_eyken_ptf_2012}.  The main novelty is
that Gaia DR2 reported a 10.3$\sigma$ ``astrometric excess'', indicating
that the residuals to the best-fitting astrometric model were larger
than expected based on the statistical uncertainties.
We comment on the significance and
interpretation of this excess in Section~\ref{sec:gaia}.

\subsubsection{Hierarchical Cluster Membership}
\label{subsec:hierarchical}

Gaia also provided astrometric parameters for tens of thousands of
young stars in the Orion complex.  Stellar populations in giant
molecular cloud complexes are not monolithic; substructured groups are
the norm \citep{briceno_lowmassOB_2007}.  The Orion molecular cloud
complex in particular has numerous subgroups, with ages ranging from 0.5
to 15$\,$Myr. For an incomplete sampling, see for instance
\citet{briceno_cida_2005,jeffries_kinematic_2006,briceno_25_2007,kounkel_apogee2_2018}
and \citet{briceno_cidaII_2019}.

\ptfo\ was initially identified as a member of the Orion$\,$OB1a
sub-association and designated CVSO\,30 by \citet{briceno_cida_2005} through combined
photometry and spectroscopy.  Later work by \citet{briceno_25_2007}
clarified that \ptfo\ was in a kinematically distinct subgroup of
Orion~OB1a, named the ``25$\,$Ori'' group after its brightest
member. They reported that the 25$\,$Ori group has an isochrone age of 7--10$\,$Myr, and
a smaller fraction of stars with disks than younger nearby sub-associations
\citep{hernandez_spitzer_ob1_2007}.

With the Gaia astrometry, it has become clear that 25$\,$Ori itself
has distinct subgroups
\citep{kounkel_apogee2_2018,briceno_cidaII_2019}.  In describing the
cluster membership of \ptfo, we follow the notation and results of
\citet{kounkel_apogee2_2018}.  These authors combined astrometric data
from Gaia DR2 with near-infrared spectra from APOGEE-2
\citep{gunn_sdss_2006,majewski_apache_2017,blanton_sloan_2017,zasowski_target_2017,cottle_apogee2_2018}.
They performed a hierarchical clustering on the six-dimensional
position and velocity information to identify subgroups within the
Orion complex.  From smallest to largest, \ptfo\ was identified
as a member of the following hierarchical subgroups:
\begin{equation}
  %{\rm PTFO\,8\text{-}8695}
  % \in
  {\rm 25\,Ori\text{-}1}
  \subset {\rm 25\,Ori}
  \subset {\rm Orion\ OB1a}
  \subset {\rm Orion\ D},
\end{equation}
where `$\subset$' means ``is a proper subset
of''.  25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149
identified members.  Based on the CMD, the mean age was found to be
6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
\citep{kounkel_apogee2_2018}. 
% JNW: What's the difference between a CMD and an HR diagram?
\citet{kounkel_apogee2_2018} identified
seven other smaller groups in the Orion complex near the Be star
25$\,$Ori. These groups received higher numbers, {\it e.g.},
25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD}
=12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).

These details concerning the group membership for one object are
cumbersome to those accustomed to the simple distinction between ``young cluster members''
and ``old field stars''.  Although all members of the Orion complex are
indeed young relative to the field, these details are essential for
assessing the photometric evidence for the binarity of \ptfo, because of
the degeneracy between stellar luminosity and age for
pre-main-sequence stars.  Having a clean sample of reference stars that are
tightly associated with \ptfo\ --- both spatially
and kinematically --- minimizes contamination
not only from field stars, but also from older and younger members of the
Orion complex.

\section{TESS Analysis}
\label{sec:tess}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f3.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Phase-folded longer and shorter-period signals.}
    {\it Top}: Longer-period signal, as in Figure~\ref{fig:splitsignal}.
    {\it Bottom}: Shorter-period signal. The reference phase is set to
    the dip.  Gray points are the 10 minute cadence
    \texttt{PDCSAP} flux.  Black points are binned to 100 points per
    period.
		\label{fig:phasefold}
	}
\end{figure*}

\subsection{Inspection}

Our initial inspection of the TESS lightcurve, in both its 2-minute
PDCSAP and 30-minute FFI forms, showed a strong sinusoidal beat signal
(Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}, top
panel). As a precursor to more detailed analysis, we calculated generalized
Lomb-Scargle periodograms using \texttt{astrobase}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
The two tallest peaks occur at 0.448\,days and 0.499\,days.
We will refer to these two periods as the ``shorter period'' $P_{\rm s}$
and the ``longer period'' $P_{\rm ell}$.
The peak at $P_{\rm \ell}$ is the taller of the two peaks.
Lower-power harmonics of both signals are also present.

% 0.14 = A1+A2
% 0.06 = A1-A2
% 2A1 = 0.20
% -> A1 = 0.1
% 2A2 = 0.08
% -> A2 = 0.04

The peak-to-peak lightcurve amplitude at maximum, when the two signals
interfere constructively, is about 14\%.  During the times of destructive interference, the peak-to-peak
amplitude is about 6\%.  Assuming the signals are mainly sinusoidal,
simple algebra tells us that the peak-to-peak amplitudes should be
about 10\% for the longer-period signal, and 4\% for the shorter-period signal.
To view the phase-folded light curves of the longer-period signal,
we subtracted the best-fitting sinusoid at the shorter period; the resulting
light curve appears smooth and nearly sinusoidal.
But after subtracting the best-fitting sinusoid at the longer period,
visual inspection of the phase-folded light curve of the shorter-period signal
revealed substructure resembling the ``dips'' seen in previous
observations. In particular, there was a $\approx$1\% dip lasting about an hour.
These initial impressions turned out to be consistent with the results of our
more complicated analysis, described below.

% JNW This doesn't seem necessary to me.
%Initial signal-processing experiments fitting out splines or sinusoids
%showed that after subtracting out the long-period signal, the
%short-period signal dominated the periodogram, and vice-versa.
%However it quickly became clear that it would be beneficial to
%simultaneously model each signal, in order to preserve the
%power at each frequency.

\subsection{Lightcurve Model}

We fitted a model to the lightcurve consisting of a linear combination of Fourier modes
with periods $P_{\rm s}$ and $P_{\rm ell}$, as well as a number of harmonics chosen as described below. To try accounting for the dips, we also added an
analytic transit model with period $P_{\rm s}$.
Symbolically, the total flux $f$ is given as
\begin{equation}
  f = f_{\rm s} + f_{\rm \ell}
  = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
\end{equation}
where $f_{\rm s}$ is the flux at the shorter period, and
$f_{\rm \ell}$ is the flux at the longer period.  Writing out the
Fourier terms explicitly,
\begin{align}
  f = &f_{\rm transit,s} + \sum_{n=1}^{N} A_n \sin(n\omega_{\rm s}t)
  + \sum_{n=1}^{N} B_n \cos(n\omega_{\rm s}t)\\
  &+ \sum_{m=1}^{M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
  + \sum_{m=1}^{M} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
\end{align}
where $N$ and $M$ are the total number of modes at the shorter and longer
periods, respectively,
% JNW: maybe N_\ell and N_s?
$A_i$ and $B_i$ the amplitudes of each
mode (which can be positive or negative), and $\omega_\ell$
and $\omega_{\rm s}$ are the angular frequencies of the longer-period
and shorter-period signals. By not including a phase parameter
in the shorter-period model, we have implicitly defined the zero point
of the phase scale. The relative phase of the longer-period model
is specified by the phase parameter $\phi_\ell$.
Since we did not know in advance how many harmonics
would be appropriate to include in the model, we considered a number of different choices for
$N$ and $M$, and used the Bayesian information criterion to select
the final model (Table~1).

As an example, one possible model consists of a transit, $N=2$
sines and cosines at the shorter period, plus $M=1$
sine and cosine at the longer period.  In this case, the free parameters would
be as follows.  The transit model parameters are the impact
parameter, the planet-to-star radius ratio, two quadratic limb
darkening parameters, the planet's orbital period (set equal to $P_{\rm s}$)
the time of a particular transit, and the mean flux.  There
would be $2N=4$ additional Fourier amplitudes at the shorter period,
plus $2M=2$ Fourier amplitudes at the longer period, and well as $P_\ell$
itself and the relative phase $\phi_\ell$.  The total number
of parameters is 14.

We implemented and fitted the models using \texttt{PyMC3}, which is
built on \texttt{theano}
\citep{salvatier_2016_PyMC3,exoplanet:theano}.
For the Fourier terms,
we used the default math operators.
% JNW what does that mean? If you just mean 'numpy', is it necessary to state?
For the exoplanet transit, we
used the model and derivatives implemented in the \texttt{exoplanet} code
\citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.  To
speed up the fitting process, we averaged the 2-minute lightcurve
to 10-minute samples.  We correspondingly scaled the uncertainties in
the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
initialized each model with the parameters of the maximum a posteriori (MAP) model.
We then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
diagnostic \citep{gelman_inference_1992}.  We tested our ability to
successfully recover injected parameters using synthetic data before
fitting the \ptfo\ lightcurves.

\subsection{Fitting Results}

We considered nine models, with the number of modes per frequency
$N$ and $M$ ranging from one to three.  To select our preferred model,
we used the Bayesian information criterion (Table~1).  The model with
the lowest BIC had two modes at the shorter 10.74$\,$hr period, and
two modes at the longer 11.96$\,$hr period.  The next-best model had
an additional harmonic at the longer period ($M=3$), but was otherwise
identical.  All nine models have reduced $\chi^2$ values ranging between 1.37
and 1.51, which suggests a plausible though imperfect agreement
between the data and the model to within the formal uncertainties.
Table~2 gives the best-fitting parameters for the preferred model,
which has the lowest BIC value.

To explore where each model succeeded and failed, we split the raw
signal into its respective components (Figures~\ref{fig:splitsignal}
and~\ref{fig:splitsignalii}).  We also examined the phase-folded
signals (Figure~\ref{fig:phasefold}).

In every model, the $11.96\,{\rm hr}$ variability is a simple sinusoid
with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$
variability is always more complex.  The overall impression is of
a distorted sinusoidal function,
with a peak-to-peak
amplitude of about 4\%.
The asymmetric sinusoid rises to a maximum near phase 0.25,
and reaches minimum brightness between phases $-0.5$ and $-0.25$.
Between phases $-0.5$ and $0.0$ there appears to be complex shorter-timescale variability, 
including an abrupt jump at phase $\approx -0.33$,
and ending with a ``dip'' 
of depth $\approx$1.2\%, lasting $\approx$0.75 hours.

The periodogram of the final residual (Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}
bottom row) shows a barely significant and poorly-resolved peak at
$\approx$8 days, consistent with the visual impression of some slower
trends in the residuals after subtracting the shorter- and longer-period
signals.

\section{Tests for Binarity}
\label{sec:gaia}

\subsection{Visual Binarity}
\label{subsec:blend}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene used for blend analysis.}
		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
		logarithnic grayscale. The yellow star is the position of \ptfo\.
		Orange crosses are neighboring stars with $T<17$. The
		\texttt{X} and \texttt{/} hatches show the apertures
		used to measure the background and target star flux, respectively.
		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
		field, with a linear grayscale. The circles show the apertures of radii
		1, 1.5, and 2.25 pixels used in our blend analysis. To the northwest
		of \ptfo\ and between the
		blue and orange circles is ``Star A'',
		the only star bright and close enough to be contributing
		to the signal attributed to \ptfo\. However, the pixel-level TESS data
		showed that any such contamination is negligible (see
		Section~\ref{subsec:blend}).
		% JNW Label Star A?
		\label{fig:scene}
	}
\end{figure}

The TESS pixels are $\approx21$'' per side. Before making any
interpretations, we needed to consider whether light from
neighboring stars could have contributed to the photometric signal
we are attributing to \ptfo. The scene is
shown in Figure~\ref{fig:scene}.  In the upper panels, the pixels used
to measure the background level in the SPOC lightcurve are indicated
with `\texttt{X}' hatching, and the pixels used in the final
lightcurve aperture are shown with `\texttt{/}' hatching.

The target star, \ptfo\ (TIC 264461976), has a $T$-band magnitude of
14.0, and its position is shown with a star.  The other (unlabeled)
star inside the target aperture, TIC 264461979, has $T=16.8$ and so
cannot contribute more than about 10\% to the total signal.  The only
neighbor that is sufficiently close and bright that its light might
contaminate the target star is TIC 264461980, with $T=14.8$, which we
dub ``Star A''.  Star A is 23.6'' northwest of the target, and based on the
magnitude difference could contribute flux variations as large as 48\% the flux of our
target star, \ptfo.

The variability of \ptfo\ with a period consistent with $P_{\rm s}$ had already been observed
based on images with arcsecond resolution.
Thus, our main concern regarding blending
was whether 
the longer-period signal with period $P_{\rm \ell}$ originated from \ptfo,
or from Star A. We took two approaches to investigate the source of the long-period
signal.

First, we examined the CDIPS FFI lightcurves of the target, which are
available on MAST \citep{bouma_cluster_2019}. Three lightcurves are available,
based on photometric apertures with a radius of 1, 1.5, or 2.5 pixels. The maximal
peak-to-peak beat amplitude was the same to within a percent,
regardless of the size of the photometric
aperture that was used to create the lightcurve.  If Star A
were the source of the long-period variability, we would expect the
peak variability amplitude to be smallest in the 1 pixel aperture,
based on the separation of the sources (Figure~\ref{fig:scene},
bottom).  From this test alone, it seems unlikely that Star A is the
source of the long-period signal.

Second, we examined the 2-minute lightcurve of each individual pixel in the scene,
using the interactive tools implemented in
\texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
source of the long-period variability, we would expect the pixels
nearest to Star A to show a sinusoidal signal with amplitude exceeding
$10\%$.  The data do not show this pattern.  The pixel directly
below Star A does not clearly show any sinusoidal variability; the
peak-to-peak variability in that pixel is $\lesssim 8\%$.  In
contrast, the south-easternmost pixel within the \ptfo\ aperture (the
pixel furthest from Star A that was used in the optimal aperture)
shows the longer-period sinusoidal variability signal with an amplitude of
14\%.  We conclude that within the resolution of the Gaia
DR2 source catalog, the $P_{\rm s}$ and $P_{\rm \ell}$ signals
originate from \ptfo.  Based on the work of \citet{ziegler_measuring_2018}, we can
surmise that stellar companions with separations wider than $\approx$1'' (349~AU) and
within $\Delta G \approx 3$ magnitudes of \ptfo\ would have likely
been detected through this approach.

\citet{van_eyken_ptf_2012} obtained stronger constraints on possible
stellar companions based on high-resolution imaging with the NIRC2 camera on Keck II.  They reported
3-$\sigma$ $H$-band magnitude difference limits of 4.3, 6.4, and 8.9 at
angular separations of 0.25, 0.5, and 1.0 arcseconds (87, 175, and
349~AU).  They also detected a point source, not present in Gaia DR2,
7.0 magnitudes fainter than the target, and 1.8$''$ to the north-east.  Due
to the brightness difference, this point source\footnote{This
point source was claimed to be a potential planetary-mass object
\citep{schmidt_direct_2016}. Subsequent analysis of its colors showed
that it is a background star \citep{lee_evidence_2018}.} cannot be the
source of our signals.

%FIXME: would be better to actually be able to get them...
While long-baseline radial velocity (RV) measurements could provide
complementary lower limits on companion masses and separations, the available RV
data for \ptfo\ have large uncertainties due to the faintness of the star.  The
RV dataset with the longest timespan in the literature appears to be the 5
Keck/HIRES measurements acquired over 10 days in April 2011 by
\citet{van_eyken_ptf_2012}.  The root-mean-squared RV over that 10 day span was
$1.6\,{\rm km}\,{\rm s}^{-1}$, consistent with the measurement
precision.  Although \citet{yu_tests_2015} acquired 22 further
Keck/HIRES spectra on one night in December 2013, those points
were not reduced to velocities. Further Keck/HIRES
measurements of this target could confirm or refute the existence of
possible binary companions.

\subsection{Photometric Binarity}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f5a.pdf}
		}

		\vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Evidence for binarity in \ptfo}.
    {\it Top:} Hertzsprung-Russell diagram of \ptfo\ and late-type
    members of 25$\,$Ori-1. Members of the 25$\,$Ori-1 group (black
    circles) were identified by \citet{kounkel_apogee2_2018}.  The
    gray circles are stars in the ``neighborhood'', i.e.,
    non-member stars within 5
    standard deviations of the mean 25$\,$Ori-1 right ascension,
    declination, and parallax.  It contains members of the Orion
    complex with its full spread of ages, in addition to field
    interlopers.  $G$ denotes Gaia broadband magnitudes, $Bp$ Gaia
    blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
    arcseconds.  The $x$-axis limits are set to show only K and
    M dwarfs, to accentuate \ptfo's separation from the
    single-star sequence.  {\it Bottom:} Astrometric goodness-of-fit
    versus $Rp$ magnitude for 25$\,$Ori-1 members.  The single-source
    astrometric model for \ptfo\ provides a poor fit, which could be due to
    stellar variability or binarity.  Cluster members that
    are at least as variable as \ptfo\ show lower astrometric excesses
    (black squares), suggesting binarity as the root cause.
		\label{fig:gaia}
	}
\end{figure*}

We also used the Gaia data to see if the observed luminosity of \ptfo
is too bright to be a single star, i.e., if it is a ``photometric
binary.''
To assemble a set of stars coeval with \ptfo,
we used the 25$\,$Ori-1 members identified by
\citet{kounkel_apogee2_2018}, and discussed in
Section~\ref{subsec:hierarchical}.

To define a set of non-member stars that nonetheless are subject
to similar selection criteria,
we defined the reference ``neighborhood'' as the
group of at most $10^4$ randomly selected non-member stars within 5
standard deviations of the mean 25$\,$Ori-1 right ascension,
declination, and parallax.  We queried Gaia DR2 for these stars using
\texttt{astroquery} \citep{astroquery_2018}.  This yielded 1{,}819
neighbors.  While some of these stars may indeed be members of the
Orion complex, or even of 25$\,$Ori-1, enforcing this cut on positions
and parallaxes ensures that we are comparing stars with similar
amounts of interstellar reddening.

We examined the resulting five-dimensional distribution of right ascension,
declination, proper motion in both directions, and parallax.  The first point we noted
was that 25$\,$Ori-1 is a clearly defined over-density in each
dimension: the cluster was confirmed to exist, and to be distinct from the neighborhood. \ptfo\ was also within the cluster in each of these projected
dimensions.

%Given our detection of two separate signals, whether \ptfo\ could be a
%photometric binary was of great interest.  
Figure~\ref{fig:gaia} shows
the HR diagram we constructed from the data.  The diagram shows
that \ptfo\ is $\approx$0.75 magnitudes brighter than the average
25$\,$Ori-1 star of the same color.  In other words, it is about twice
as bright as expected for a single star in the cluster.
It also seems to be part of a ``photometric binary''
track that runs parallel to the main track.

The implication is that either {\it (i)} \ptfo\ is notably younger
than the kinematically identical 25$\,$Ori-1 members, or {\it (ii)}
\ptfo\ is a binary with two components of nearly equal brightness.
Since there is no other reason to
suspect an age difference, and because the source showed two separate
photometric signals with similar but distinct periods,
the binary interpretation seems more probable.

\subsection{Astrometric Binarity}

A separate possible line of evidence for binarity is the Gaia DR2
astrometry.  As noted in Section~\ref{sec:observations}, the Gaia DR2
astrometric solution for \ptfo\ shows a 10.3$\sigma$ ``astrometric excess'',
a parameter that quantifies the degree to which a single-star model
fails to fit the astrometric measurements.  Specifically,
the single-source astrometric model yielded $\chi^2=325.2$.  There are
121 astrometric measurements, and 5 free parameters, and therefore 116
degrees of freedom. The reduced $\chi^2$ is 2.80.  The majority of
stars with comparable brightness in Gaia do not show such poor
goodness-of-fit \citep[see][Appendix A]{lindegren_gaiasoln_2018}.

Potential explanations for the poor astrometric fit include
photometric variability and unresolved stellar binarity \citep[{\it
e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.  If
photometric variability were the cause, we would expect comparably
faint stars in the same kinematic group of Orion to show similar
astrometric excesses, as the majority of young stars are highly
variable.

% NB: CDIPS-I only made 131 light-curves for 25Ori-1, because I wasn't
% aware of the Kounkel+18 member list at the time.
Using the same 149 members in the 25$\,$Ori-1 subgroup, we calculated
the astrometric reduced $\chi^2$ for each member.  We then queried the
CDIPS lightcurve database at MAST \citep{bouma_cluster_2019} to find
the subset of members that were at least as variable as \ptfo.
We measured the variability amplitude by taking the difference between the
$95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux measurements.
This yielded 30 stars of equal or greater variability.
The lower panel of Figure~\ref{fig:gaia} shows the
reduced $\chi^2$ as a function of stellar brightness.  \ptfo\ is in
the upper 90$^{\rm th}$ percentile of stars showing astrometric
excesses within the 25$\,$Ori-1 group.  Relative to other M-dwarf
group members with comparable brightnesses and variability
characteristics, \ptfo\ still stands out by virtue of its failure
to conform to a single-star astrometric model. This supports
the interpretation that \ptfo\ is a binary star.

We will have to wait for the full release of the
nominal Gaia mission to determine definitively whether the astrometric
excess is caused by stellar binarity or photometric variability.
Nonetheless the fact that comparably variable stars do not show
comparably large astrometric excesses suggests that stellar binarity
is indeed the root cause.

\section{Discussion}
\label{sec:discussion}

\subsection{Longer-Period Signal}

The standard interpretation for 11.96$\,$hour nearly sinusoidal modulations
of a pre-main-sequence M dwarf is stellar rotation.  This is the
dominant signal in the system with 10\% amplitude, and there is no
evidence to suggest that this signal has any other origin.

In their report on the discovery of the unusual
photometric variability, \citet{van_eyken_ptf_2012} saw an alias of the
same signal ({\it e.g.}, their Figure~7),
% JNW: Don't understand this sentence. What is the 'same signal'?
and identified it as a
periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it to
their observing cadence, because of its close correspondence to the
sidereal day.
% JNW What is this next phrase doing here?
%While the TESS data can show significant reflected
%light from the Earth \citep[{\it e.g.},][]{luger_tess_2019}, 
Our
pixel-level analysis showed that the signal is specific to only pixels
near \ptfo, and no other pixels.  We therefore conclude that the
signal is not an artifact of systematic errors.

We are not the first to reach the conclusion that the long period
sinusoidal modulation is astrophysical.  A study by
\citet{koen_multicolour_2015} identified the same modes and aliases as
\citet{van_eyken_ptf_2012}, but argued that the signal was
astrophysical (however they were still unsure of the exact period).
Using photometry from the YETI global telescope network,
\citet{raetz_yeti_2016} eventually came to the conclusion that the
$0.50\,{\rm d}$ signal was indeed from stellar rotation.  The TESS
data strongly support this conclusion.

\subsection{Shorter-Period Signal, Including the ``Dip''}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f6.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Timing residuals for \ptfob\ based on a decade of monitoring.}
		Black points are times of dips, minus the indicated linear
		ephemeris.  The phase of the shorter-period signal
		is plotted on the $y$-axis. The star symbol represents
		the TESS ephemeris.
		Dips were observed by \citet{van_eyken_ptf_2012},
		\citet{ciardi_followup_2015}, \citet{yu_tests_2015},
		\citet{raetz_yeti_2016}, \citet{onitsuka_multicolor_2017}, and
		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
		at phase 0 in mid-2019) are consistent with noise, and were likely
		reported because dips were expected, rather than
		convincingly observed.  Horizontal dashed lines are drawn at
		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting either a
		numerical coincidence or an observational bias.  The orbital phase
		observed by TESS (lower-right) is consistent with that of
		\citet{tanimoto_evidence_2020}.
		\label{fig:o_minus_c}
	}
\end{figure}

The TESS lightcurve shows a dip that lasts about 45 minutes, and seems
to recur every 10.74 hours
(Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
The dip duration is roughly the same as that observed by previous
investigators \citep{van_eyken_ptf_2012,yu_tests_2015}.  The 1.2\%
depth is similar to what has been observed in the near-infrared
\citep{onitsuka_multicolor_2017}.  However the dip depth seems likely
to have evolved over time between being not present at all, to a
maximum of $\approx$5\% \citep[{\it
e.g.},]{koen_multicolour_2015,yu_tests_2015,tanimoto_evidence_2020}.

One particularly interesting feature of the dip is its epoch.  The dips do
not occur with strict periodicity
\citep{yu_tests_2015}.  In fact, \citet{tanimoto_evidence_2020}
provided stark evidence for different behavior altogether:
over a time-span of years, the dip ``split'' into distinct groups at
particular repeating phases.  See for instance their Figures~2
through~4.  Fitting a decade of observations, they provided the
following linear ephemeris, which we did not find any need to update.
\begin{align}
t_0\ {\rm BJD}_{\rm TDB} &= 2455543.943 \pm 0.002 \\
P &= 0.4483993 \pm 0.0000006 d.
\end{align}

Figure~\ref{fig:o_minus_c} shows the phase of the dip we detected
in the TESS data, relative to their ephemeris.  It agrees with the
independent December 2018 measurements by
\citet{tanimoto_evidence_2020}: either the dip abruptly shifted phase
over the past decade or, more likely, there are multiple dips that
have come and gone at different phases.

Figure~\ref{fig:o_minus_c} shows two additional strange features: {\it
(i)}  multiple dips per cycle, and {\it (ii)} a set of dips
numerically coincident with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
s}$.  The observation of multiple dips per cycle in 2015 was seen
independently by both \citet{yu_tests_2015} and
\citet{tanimoto_evidence_2020}.  It therefore seems credible.
Inspecting the \citet{tanimoto_evidence_2020} lightcurves, the claim
of multiple dips per cycle in December 2018 at phase 0 and $-0{.}47$
seems less plausible---the phase $-0{.}47$ dips are strongly detected,
while the suggested phase 0 dip is not clearly present in the data.

We are not sure what to make of the numerical coincidence.  The ratio
of long to short periods is roughly 10:9.  It is not clear that this
would obviously translate into an observational bias unless by some
fluke three season's worth of observations managed to only observe
every ninth dip.  This is of course not the case, and we therefore
leave this curiosity as observation {\it sans} interpretation.
% JNW: This seems like an important clue. I don't have the mental
% energy to ponder it now, but it deserves pondering.

\subsection{Short Period Out-of-Dip Modulation}

Visually, the out-of-dip modulation at the 10.74$\,$hour period
resembles a slightly asymmetric sinusoid (Figure~\ref{fig:phasefold}).
Non-zero contributions in both the first and second harmonics are
detected (Table~2).  The first sine and cosine harmonic both have
amplitudes of roughly $0.9\pm0.1\%$.  The second sine harmonic has
amplitude $0.16 \pm 0.07\%$, so is non-zero at a significance of only
2.3$\sigma$.  The second cosine harmonic has negative amplitude $0.53
\pm 0.06\%$.  In our sign convention, the fact that it is negative
means that this component peaks at phase 0.25 and 0.75.

\subsubsection{Ellipsoidal Variability?}
If there were a giant planet transiting \ptfo, it would tidally
distort the host star, and cause ellipsoidal photometric modulations
that also peak at quadrature \citep[see][]{shporer_astrophysics_2017}.
Interpreting the second cosine harmonic as planet-induced tidal distortion,
it would imply a minimum planet mass $M_{\rm p} \sin i$ of
$3.8\,M_{\rm Jup}$.  For this estimate, we assumed $R_\star = 1.39
R_\odot$, and $M_\star = 0.39 M_\odot$ \citep{van_eyken_ptf_2012}.
This ellipsoidal amplitude is larger than the typical
modulations induced by close-in giant planets because the host star is
puffy, and still on the pre-main-sequence.

The planetary interpretation however does not readily explain the
first sine and cosine harmonics.  Interpreting the sine component as
Doppler beaming would imply a secondary mass greater than the primary
($0.86\,M_\odot$).  Interpreting the cosine component as reflected or
emitted light from the planet's surface is nonsensical because the sign is wrong---the
planet would need to be {\it absorbing} light.

\subsubsection{Similar Lightcurves}
\label{subsec:dipstars}

\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf \ptfo\ and its brethren.}
    Five transient and persistent flux dip stars selected based on
    their visual similarity to the short-period signal in \ptfo are as
    shown.
    They include
    EPIC 204143627,
    EPIC 204270520,
    EPIC 204321142,
    EPIC 246938594,
    and
    EPIC 205483258 (RIK-210).
    %and EPIC 204787516.
    RIK-210 has the longest period of any of these objects. 
    All the analogs displayed are either in Taurus or Upper~Sco, and
    meet the characteristics of Section~\ref{subsec:dipstars}.
    % EPIC 204787516 has a less angular dip, and could also be an
    % eclipsing binary.
    % All stars have similar properties as \ptfo, but are in Upper~Sco.
    We found these objects through studies by
    \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and
    \citet{rebull_usco_2018}.
		\label{fig:brethren}
	}
\end{figure*}

When physical explanations are not forthcoming, taxonomy is a
useful exercise.  By searching the literature, we have found about a dozen lightcurves with
similar morphologies to \ptfo, drawn from surveys of low-mass weak-lined
T Tauri stars in regions including $\rho$ Oph, Upper~Sco, Taurus, and
perhaps the Pleiades
\citep{rebull_rotation_2016,david_transient_2017,stauffer_orbiting_2017,stauffer_rotevol_2018,rebull_usco_2018,rebull_rotation_2020}.
These surveys were performed using K2 \citep{howell_k2_2014}.  We
downloaded some of these lightcurves from MAST, opting for the
EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.
They are plotted in Figure~\ref{fig:brethren}.

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
In the terminology of \citet{stauffer_orbiting_2017},
these objects are morphologically distinct from ``scallop shell'' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars all show angular dips
that are cannot be explained as the effects of starspots.
These stars typically have the following things in common:
\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
  \item They are weak-lined T Tauri stars.
  \item The spectral type is M2 to M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item The age is $\lesssim$ 100$\,$Myr.
%  \item They lack spectroscopic indicators of strong accretion.
% JNW: Same as WTTS?
  \item The lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item The rotation is rapid---usually between 0.5 and 2.0 days.
  \item There is rarely any detectable infrared excess in WISE data
  (never any W4 detection; only a few W3 detections).
  \item They sometimes show multiple dips per cycle.
  %({\it e.g.}, EPIC 203692610, RIK-210).
  \item The dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item The dip depths can change after flares.
  \item They are rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

There are two crucial additional points concerning the transient flux
dips.  First, the dip durations seem to scale linearly with
the photometric periods \citep[][Figure~26]{stauffer_orbiting_2017}.  In the usual
idealized limits, the transit duration $T$ of a point source across
the stellar disk scales as $T \propto R_\star (P/M_\star)^{1/3}$
\citep{winn_exoplanet_2010}.  While the shortest period
$\approx$0.5-day transient flux dip stars have dip durations
consistent with point sources, at longer periods of 1 to 5 days the
dip durations become many hours, far too long to be caused by
planetary transits.

Second, between 40-50\% of the transient flux dip stars discovered in
$\rho$~Oph and Upper~Sco show two
Lomb-Scargle periods, and so are apparently binaries
\citep[][Table~1]{stauffer_orbiting_2017}.
This is higher than the main-sequence companion
fraction of ${\rm CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
\citep{henry_solar_2006,duchene_stellar_2013}.
Low-mass pre-main-sequence stars however have been shown to 
companion fractions up to twice as high in dispersed clusters such as
Upper~Sco and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.
A detailed high-resolution imaging survey would be necessary to determine
whether the transient flux dip stars truly have any distinct
population-level binarity properties relative to other young low-mass
stars.

\subsection{Physical Interpretation}
\label{subsec:physical}

The evidence for binarity in \ptfo\ is as follows.  First, the star is
photometrically twice as bright as stars of the same color in its
kinematic group (Figure~\ref{fig:gaia}).  Second, it shows two distinct
photometric signals.  These points alone suggest binarity
\citep{stauffer_rotevol_2018}.  For the case of \ptfo, there is a
third line of evidence: the Gaia DR2 entry for \ptfo\ reports a poor
fit of the single-star model to the astrometric data.  While this could be caused
by stellar variability, other cluster members
that are just as variable do not typically show the same level of
excess astrometric motion.
Therefore the astrometric excess is a suggestive third line
of evidence for binarity in \ptfo. To us, the evidence leads to
the conclusion that \ptfo\ is a nearly equal-mass binary consisting
of two rapidly rotating stars.

Based on the lack of an infrared excess seen by \citet{yu_tests_2015},
the primordial gas disks seem to be have been depleted\footnote{A
potential confounding factor is completeness. Is Spitzer sensitive to
infrared excesses at the distance of Orion? The series of studies by
\cite{hernandez_spitzer_2006,hernandez_spitzer_ob1_2007,hernandez_spitzer_sig_2007,hernandez_spitzer_2009}
show that it is.} around both stars in \ptfo. The stars are therefore
no longer magnetically locked to their disks.  This is consistent with
the $\approx$half-day periodicities of both rotation signals: young
disked M dwarfs typically rotate with periods of two days or more due
to magnetic locking \citep[{\it e.g.},][]{rebull_rotation_2020}.  If
the two stars are within $\approx$50$\,$AU of each other, as required
by the NIRC2 adaptive optics imaging, then it would also be expected
that the stars would truncate the outer edges of their respective
disks, in a manner seen at the population level in exoplanetary
systems \citep{kraus_impact_2016,moe_impact_2019}.  This hard outer
boundary condition could propagate to the inner disk and affect its
evolution.
% JNW I wonder if there could be a circumbinary ring, a la KH 15D.

The main physical question is what causes the
transient dips. This is an unsolved problem not only for \ptfo\ but
also for an emerging class of similar young rapidly rotating M-dwarfs.
Many possible explanations been detailed by \citet{rebull_rotation_2016}, \citet{david_transient_2017},
\citet{stauffer_orbiting_2017}, and \citet{zhan_complex_2019}.
Their disfavored explanations include that the dips are caused by
{\it (i)} eclipsing binaries;
{\it (ii)} ``dipper''-flavor Class-I or Class-II disks;
{\it (iii)} eclipses of prominences;
{\it (iv)} high-latitude accretion hotspots;
{\it (v)} high-latitude starspots;
or
{\it (vi)} dust clouds of plausible composition.
We also view the possibility of {\it (vii)} tidally disrupted planetary
or cometary material to be implausible, given the synchronicity
between dip and rotation periods seen across many systems.

The explanations that are not yet ruled out include
{\it (i)} transiting clumps of gas at the Keplerian corotation radius;
{\it (ii)} transits of enshrouded protoplanets;
{\it (iii)} occultations of starspots by an optically thick disk.
The first and last explanations have added appeal because they are
flexible enough to explain not only the transient and persistent-dip
M-dwarfs, but also the ``scallop shell'' M-dwarfs
\citep{stauffer_orbiting_2017}.
Despite this appeal, the possibility of distinct mechanisms explaining
these distinct variability classes remains open.

The evolution of \ptfo\
over the past decade (Figure~\ref{fig:o_minus_c}) could offer important
hints.  Specifically, \ptfo's transition between having none, one, and
multiple dips per cycle seems important.  It strains the ``enshrouded
protoplanet'' interpretation, because there are no known processes
that cause a planet's orbital phase to jump.  The dips would then need
to be caused by material that was somehow disrupted from the planet,
but somehow remained co-orbital for an extended duration. This seems
implausible.

\section{Conclusions}
\label{sec:conclusions}

The combination of TESS and Gaia data has clarified a few things about
the \ptfo\ system.  Our main results are as follows.
\begin{itemize}
  \item {\it The TESS light-curve shows two periodic signals.} The
    ``long'' signal is a 10\% peak-to-peak sinusoid that repeats every
    11.96 hours.  The ``short'' signal is a 4\% peak-to-peak ``dip +
    asymmetric sinusoid'' that repeats every 10.74 hours. The signals
    beat, and therefore cannot be an artifact linked to data
    processing.
    Within the angular resolution of the Gaia source
    catalog, both signals need to originate from \ptfo.
  \item {\it The Gaia data imply binarity.} Relative to stars in its
    kinematic group, \ptfo\ is a photometric binary
    (Figure~\ref{fig:gaia}, top).  Relative to stars in its group that
    are at least as photometrically variable, \ptfo\ also shows
    signs of astrometric binarity (Figure~\ref{fig:gaia}, bottom).
  \item {\it The orbital phase of the dip has changed since the
    discovery by \citet{van_eyken_ptf_2012}.} As shown in
    Figure~\ref{fig:o_minus_c}, the phase seems to have jumped,
    perhaps twice. This agrees with the recent study by
    \citet{tanimoto_evidence_2020}.
  \item {\it All properties of \ptfo\ are consistent with the emerging
    class of transient and persistent flux dip stars.} Analogous
    lightcurves are shown in Figure~\ref{fig:brethren}.  Properties of
    this variability class are enumerated in
    Section~\ref{subsec:dipstars}.
\end{itemize}

The physical mechanism that explains the transient and persistent flux
dips is unresolved. Our preferred explanations include
transiting clumps of gas at the Keplerian corotation radius, and
occultations of starspots by an optically thick disk \citep[{\it
e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
The jumping orbital phase disfavors the explanation of an
enshrouded, transiting protoplanet.  Though \ptfob\ may not be a planet,
as we and others had hoped, understanding \ptfo\ and its analogs is a worthy
problem that might even teach us about the natal environments of the
majority of habitable-zone Earth-sized planets in the Milky Way
\citep{dressing_occurrence_2013}.
% JNW: seems like a stretch but I will not strike it down

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgements
%
%This paper includes data collected by the TESS mission, which are
%publicly available from the Mikulski Archive for Space Telescopes
%(MAST).
%
%Funding for the TESS mission is provided by NASA's Science Mission
%directorate.
%
The authors thank D.~Fabrycky for helpful tidal evolution calculations
that were eventually omitted.
%
We thank the Heising-Simons Foundation for their generous support of this work.
%
\ptfo\ was included on the TESS ``short-cadence'' target list thanks to the
Guest Investigator programs of S.\ Czesla and C.\ Huang (G011128 and
G011132 respectively).
%
The Digitized Sky Survey was produced at the Space Telescope Science
Institute under U.S. Government grant NAG W-2166.
Figure~\ref{fig:scene} is based on photographic data obtained using
the Oschin Schmidt Telescope on Palomar Mountain.
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019}
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:agol19}
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano}.
  \texttt{IPython} \citep{perez_2007},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
  \texttt{tesscut} \citep{brasseur_astrocut_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
 	{\it Imaging}:
  Second Generation Digitized Sky Survey,
 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	{\it Spectroscopy}:
 	Keck:I~(HIRES; \citealt{vogt_hires_1994}).
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
% 	El Sauce:0.356m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
 	TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}

%
% The following are entries from Table 1 that are not otherwise cited
% in the text
%
% \nocite{wilson_wasp-4b_2008}
% \nocite{gillon_improved_2009}
% \nocite{winn_transit_2009}
% \nocite{hoyer_tramos_2013}
% \nocite{dragomir_terms_2011}
% \nocite{sanchis-ojeda_starspots_2011}
% \nocite{nikolov_wasp-4b_2012}
% \nocite{ranjan_atmospheric_2014}
% \nocite{huitson_gemini_2017}

% \input{WASP-4b_transit_time_table.tex}
% \input{WASP-4b_rv_table.tex}
% \input{model_fit_table.tex}
% \input{rv_model_posterior_table.tex}
% \input{pdot_table.tex}

% \clearpage
\input{model_comparison_table.tex}
\input{model_posterior_table.tex}
% \clearpage

\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\listofchanges

\end{document}