### 2020-06-12 paper_review/response_20200605.txt
WIP for the referee report
added some nice details to flesh out the torus of gas+dust in favor of
the starspot occultations

Dear Dr. Shafter,

We thank you for organizing the review of our paper, and also
thank the referee for their helpful comments.  We have
considered the referee's remarks carefully and revised our
manuscript accordingly.

Below, we have reproduced the relevant portions of the
referee's report, alongside our responses. A list of changes is
appended to the manuscript using the trackchanges AASTeX
macros.

Sincerely,

Luke Bouma

******************************************
REFEREE COMMENT
----------
Overall: It would help if the authors used hours or days throughout the paper
in a consistent manner. The period is referred to as 0.5d and 11.98h just a
couple paragraphs apart (e.g. in Section 5.1). The abstract uses hours, most
plots use days.

RESPONSE
----------
> Throughout the manuscript, we have shifted the units wherever possible to
> hours.
>
> Exceptions include the ephemeris (which is quoted in days), and the periods
> listed in Figures 3 and 6, which are quoted in days. The former is standard
> convention. The latter was because we felt it important to list the periods
> used to generate the figures at a higher precision, and quoting higher
> precision in units of hours seems atypical in a field in which days are more
> typically seen.

******************************************
REFEREE COMMENT
----------

Introduction: The introduction takes a light/jovial tone, e.g., did they
actually wish it were a planet, or is it more that they aimed to confirm it and
found the opposite? I would normally suggest making more formal. However, I
enjoyed reading it. I'm sorry the authors ruined their own favorite planet.

RESPONSE
----------
> So are we!  Our aim was to clarify what could be happening in the system, and
> hopefully this contribution is valuable for that.  When a planetary door
> closes, a window on young star variability opens...

******************************************
REFEREE COMMENT
----------

Introduction: I do not think this is youngest hot Jupiter. That should be true
just for transiting systems, and it was the youngest at discovery, but the two
RV planets from ESPaDOnS are younger (Donati et al. 2016 and Yu et al. 2017).
Both are hot Jupiters and in Taurus. I believe the age of Tap 26 is more
ambiguous due to age spreads (multiple groups?) in Taurus, but the age of V830
Tau is certainly <3 Myr.

RESPONSE
----------

> Thank you for this correction. We have qualified the relevant sentence as the
> youngest known *transiting* hot Jupiter. We have opted to not cite the V830
> Tau or TAP 26 results, which we hope to see confirmed by independent groups.

******************************************
REFEREE COMMENT
----------

Introduction: I also think it wasn't the first H-alpha detection. That belongs
to LkCa15, which was imaged in H-alpha in 2015 (Sallum et al. 2015). Although
that planet might not exist, so this comment is probably OK. It would still be
the first close-in planet either way. I'd suggest weakening the statement
slightly to err on the side of caution.

RESPONSE
----------
> We have changed the wording to err on the side of caution, and qualify it as
> the first direct detection of Hα emission from a close-in planet.

******************************************
REFEREE COMMENT
----------

Introduction: Barnes et al. (2013) claimed that the planet and star precessed
on a period of 300-600 days, and that this could somewhat fit variable depths
and missing transits. The Ciardi et al. 2015 paper uses this as their model, in
addition to gravity darkening as mentioned in the current introduction. I
suggest mentioning the precession of the planet in the introduction, as many
papers used the precession model or at least discussed it. The line "...the
gravity-darkened star precessed..." seems to allude to this, but it's unclear
from the rest of the text.

RESPONSE
----------
> We have changed the wording of the last sentence of the first paragraph in
> the introduction to explicitly use the word "precess", rather than "torque".

******************************************
REFEREE COMMENT
----------

Section 2: The first paragraph discusses the full frame images, while the next
section discusses the 2m data. Why did the authors use the FFIs when they have
the (higher quality) 2m data?

RESPONSE
----------
> For the essential analysis of the study (i.e., extracting and fitting the
> long and short signals), we used the 2m SPOC lightcurve.
>
> We primarily mention our use of the FFI lightcurves in Section 2 because of
> the systematic jump mentioned at BJD 2458488.3, which was eventually excluded
> from further analysis. We wanted to be sure though that this jump was truly
> an instrumental systematic, and not a software systematic. Checking
> lightcurves from independent pipelines meets this goal.
>
> A secondary reason to introduce the concept of FFIs in Section 2 is that they
> are used to measure the variability amplitudes of the other members of 25
> Ori-1 in Section 4.3.

******************************************
REFEREE COMMENT
----------

Section 2.2.2: "These details concerning the group membership for one object
may seem excessive to those accustomed to the simple distinction between "young
cluster members" and "old field stars"." I suggest removing this. It's not
adding anything and sounds like the authors are making assumptions about the
reader. The membership of this object is essential for an age assessment, and
can/should be updated given the new Gaia data. Since most of the object class
in question are at similar ages, this directly relates to the study. If it had
turned out to be ~100 Myr, we would all be more skeptical (although ~100 Myr
stars like this are known, they are more rare).

RESPONSE
----------
> We've removed it -- thank you for the suggestion.

******************************************
REFEREE COMMENT
----------

Section 3/4: This deals mostly with identification of the second star. However,
the second star appears to be an assumption of section 3.2. If the periods come
from two stars that could be separated (e.g., wide binaries), it would have
been best to do so rather than try to model both by including fs and fl terms.
So it is odd that the binary search is after the light curve model in the
paper. It might make more narrative sense to move the light curve model to
after section 4, since 4 motivates 3.2.

RESPONSE
----------
> While we see the referee's point, we prefer the narrative structure of
> presenting the TESS data first, and getting into considerations of stellar
> binarity, both resolved and unresolved, afterwards.
>
> The reasons are that i) we wrote the paper primarily because of the TESS
> data; ii) the multiple signals come from TESS data. From a story-telling
> perspective, why would we worry about binarity if we didn't already know
> there were multiple signals?

******************************************
REFEREE COMMENT
----------

Section 4.3: My biggest concern with this analysis is reliance on the chi^2
from Gaia to assess binarity. This is not the suggested path from Gaia, as it
is subject to a noise model Gaia now knows has issues. For example, nearly 100%
of mid/late M dwarfs have a high chi^2 (and a large excess astrometric noise),
but clearly they are not all binaries. This is (partially) because of a
color-term in the PSF model that was not initially understood. Similarly,
almost all bright stars have a large RUWE, as do the bluest stars. The authors
correctly use RUWE, the suggested method, but the RUWE analysis is less
convincing and clearly the authors are leaning on the chi^2.

A good reference on this would be Belokurov et al. 2020, which shows what kinds
of binaries RUWE can and cannot detect. This binary might simply be too tight
to create an astrometric signal (in which case the RUWE makes sense). The
typical cutoff for binarity is a RUWE of ~1.4, see Rizzuto et al. 2018, Ziegler
et al. 2019, Pearce et al. 2020, among many others.  I suggest removing the
chi^2 analysis. If the authors want to keep it, they need to argue why it is a
more meaningful metric than RUWE, and attempt to account for color and other
parameters that impact chi^2.

RESPONSE
----------
> We've done as suggested, and replaced this sub-plot with RUWE instead of
> reduced chi^2. The text in Section 4.3 and captions have been updated
> accordingly, and chi^2 is now mostly ignored.

******************************************
REFEREE COMMENT
----------

Section 4.4:
Need a period between velocities and Further.

RESPONSE
----------
> Corrected, thank you.

******************************************
REFEREE COMMENT
----------

Section 5.2: That clear offset at (Pl-Ps)/Ps doesn't seem like a coincidence,
and the paper reads like the authors agree, but couldn't explain it. I'm not
sure I can help much here, although it is interesting. Could this be that both
stars are dippers and one turns off while the other turns on by coincidence?
Spots and hot-spot clusters can shift in phase as the magnetic fields realign,
which creates a new spot cluster at the same period but a different phase. If
it is circumstellar material, a tight binary could be shepherding the material,
but I think the systems would have to be very tight for that, and it could not
last for such a long time.

I think this can be assessed statistically; given the scatter in those points,
the odds of lining up to any given line is small. But without a physical
explanation this is hard to test. Probably this requires looking at many
dippers as an ensemble. It is not required for the primary analysis of this
paper, so I will defer to the authors if they want to dig deeper on this.

RESPONSE
----------
> We're not sure whether or not it is a coincidence, but it seemed interesting
> enough to mention.
>
> ..

******************************************
REFEREE COMMENT
----------

Section 5.3.2: The difference between the fraction of transient flux dip stars
and the binary fraction is more significant than it appears to the authors.
High mass ratio binaries would not be included, as they are too faint to detect
the second period. Further, at the distance of Upper Sco (~150pc) and assuming
Stauffer et al. 2017 would notice/separate two stars wider than a K2 pixel
(4"), the presence of two periods would not include binaries wider than 600pc.
Eyeballing the Winters et al. 2019 distribution suggests ~20% of binaries are
beyond this. Stauffer et al. 2017 also probably checked against existing
imaging data for obvious binaries down to 1-2", so likely even more are getting
cut.

RESPONSE
----------
>

******************************************
REFEREE COMMENT
----------

Section 5.4: The authors conclude this subsection with a reasonable argument
that this isn't a protoplanet. Does this same argument not rule out gas/debris
at the Keplerian coronation radius? What could maintain such clouds for
years/decades and then have them jump phase but at the same period? It's hard
to do this if it is not somehow physically attached to the star in some what
(e.g., spots can do this).

RESPONSE
----------
>

******************************************
ADDITIONAL CHANGES

A few additional changes are noted in the "List of Changes" appended at the end
of the manuscript.

We modified the text throughout to clarify that Stauffer+ 2017 favored the
explanation of a clumpy torus of gas and dust at the Keplerian co-rotation
radius, rather than our initial (incomplete) description of "clouds of gas" at
the same location.

We also modified the labels of the axes in Figure 5a for clarity.