### 2020-06-12 paper_review/response_20200605.txt
fix column widths

Dear Dr. Shafter,

We thank you for organizing the review of our paper, and also
thank the referee for their helpful comments.  We have considered
the referee's remarks carefully and revised our manuscript
accordingly.

Below, we have reproduced the relevant portions of the referee's
report, alongside our responses. A list of changes is appended to
the manuscript using the trackchanges AASTeX macros.

Sincerely,

Luke Bouma

******************************************
REFEREE COMMENT
----------
Overall: It would help if the authors used hours or days
throughout the paper in a consistent manner. The period is
referred to as 0.5d and 11.98h just a couple paragraphs apart
(e.g. in Section 5.1). The abstract uses hours, most plots use
days.

RESPONSE
----------
> Throughout the manuscript, we have shifted the units wherever
> possible to hours.
>
> Exceptions include the ephemeris (which is quoted in days), and
> the periods listed in Figures 3 and 6, which are quoted in
> days. The former is standard convention. The latter was because
> we felt it important to list the periods used to generate the
> figures at a higher precision, and quoting higher precision in
> units of hours seems atypical in a field in which days are more
> typically seen.

******************************************
REFEREE COMMENT
----------

Introduction: The introduction takes a light/jovial tone, e.g.,
did they actually wish it were a planet, or is it more that they
aimed to confirm it and found the opposite? I would normally
suggest making more formal. However, I enjoyed reading it. I'm
sorry the authors ruined their own favorite planet.

RESPONSE
----------
> So are we!  Our aim was to clarify what could be happening in
> the system, and hopefully this contribution is valuable for
> that.  When a planetary door closes, a window on young star
> variability opens...

******************************************
REFEREE COMMENT
----------

Introduction: I do not think this is youngest hot Jupiter. That
should be true just for transiting systems, and it was the
youngest at discovery, but the two RV planets from ESPaDOnS are
younger (Donati et al. 2016 and Yu et al. 2017).  Both are hot
Jupiters and in Taurus. I believe the age of Tap 26 is more
ambiguous due to age spreads (multiple groups?) in Taurus, but
the age of V830 Tau is certainly <3 Myr.

RESPONSE
----------

> Thank you for this correction. We have qualified the relevant
> sentence as the youngest known *transiting* hot Jupiter, and
> added a footnote to the V830 Tau or TAP 26 results.

******************************************
REFEREE COMMENT
----------

Introduction: I also think it wasn't the first H-alpha detection.
That belongs to LkCa15, which was imaged in H-alpha in 2015
(Sallum et al. 2015). Although that planet might not exist, so
this comment is probably OK. It would still be the first close-in
planet either way. I'd suggest weakening the statement slightly
to err on the side of caution.

RESPONSE
----------
> We have changed the wording to err on the side of caution, and
> qualify it as the first direct detection of Hα emission from a
> close-in planet.

******************************************
REFEREE COMMENT
----------

Introduction: Barnes et al. (2013) claimed that the planet and
star precessed on a period of 300-600 days, and that this could
somewhat fit variable depths and missing transits. The Ciardi et
al. 2015 paper uses this as their model, in addition to gravity
darkening as mentioned in the current introduction. I suggest
mentioning the precession of the planet in the introduction, as
many papers used the precession model or at least discussed it.
The line "...the gravity-darkened star precessed..." seems to
allude to this, but it's unclear from the rest of the text.

RESPONSE
----------
> We have changed the wording of the last sentence of the first
> paragraph in the introduction to explicitly use the word
> "precess", rather than "torque".

******************************************
REFEREE COMMENT
----------

Section 2: The first paragraph discusses the full frame images,
while the next section discusses the 2m data. Why did the authors
use the FFIs when they have the (higher quality) 2m data?

RESPONSE
----------
> For the essential analysis of the study (i.e., extracting and
> fitting the long and short signals), we used the 2m SPOC
> lightcurve.
>
> We primarily mention our use of the FFI lightcurves in Section
> 2 because of the systematic jump mentioned at BJD 2458488.3,
> which was eventually excluded from further analysis. We wanted
> to be sure though that this jump was truly an instrumental
> systematic, and not a software systematic. Checking lightcurves
> from independent pipelines meets this goal.
>
> A secondary reason to introduce the concept of FFIs in Section
> 2 is that they are used to measure the variability amplitudes
> of the other members of 25 Ori-1 in Section 4.3.

******************************************
REFEREE COMMENT
----------

Section 2.2.2: "These details concerning the group membership for
one object may seem excessive to those accustomed to the simple
distinction between "young cluster members" and "old field
stars"." I suggest removing this. It's not adding anything and
sounds like the authors are making assumptions about the reader.
The membership of this object is essential for an age assessment,
and can/should be updated given the new Gaia data. Since most of
the object class in question are at similar ages, this directly
relates to the study. If it had turned out to be ~100 Myr, we
would all be more skeptical (although ~100 Myr stars like this
are known, they are more rare).

RESPONSE
----------
> We've removed it -- thank you for the suggestion.

******************************************
REFEREE COMMENT
----------

Section 3/4: This deals mostly with identification of the second
star. However, the second star appears to be an assumption of
section 3.2. If the periods come from two stars that could be
separated (e.g., wide binaries), it would have been best to do so
rather than try to model both by including fs and fl terms.  So
it is odd that the binary search is after the light curve model
in the paper. It might make more narrative sense to move the
light curve model to after section 4, since 4 motivates 3.2.

RESPONSE
----------
> While we see the referee's point, we prefer the narrative
> structure of presenting the TESS data first, and getting into
> considerations of stellar binarity, both resolved and
> unresolved, afterwards.
>
> The reason is that the multiple signals come from TESS data
> processed at the ``default'' level. From a story-telling
> perspective, why would we worry about binarity if we didn't
> already know there were multiple photometric signals?

******************************************
REFEREE COMMENT
----------

Section 4.3: My biggest concern with this analysis is reliance on
the chi^2 from Gaia to assess binarity. This is not the suggested
path from Gaia, as it is subject to a noise model Gaia now knows
has issues. For example, nearly 100% of mid/late M dwarfs have a
high chi^2 (and a large excess astrometric noise), but clearly
they are not all binaries. This is (partially) because of a
color-term in the PSF model that was not initially understood.
Similarly, almost all bright stars have a large RUWE, as do the
bluest stars. The authors correctly use RUWE, the suggested
method, but the RUWE analysis is less convincing and clearly the
authors are leaning on the chi^2.

A good reference on this would be Belokurov et al. 2020, which
shows what kinds of binaries RUWE can and cannot detect. This
binary might simply be too tight to create an astrometric signal
(in which case the RUWE makes sense). The typical cutoff for
binarity is a RUWE of ~1.4, see Rizzuto et al. 2018, Ziegler et
al. 2019, Pearce et al. 2020, among many others.  I suggest
removing the chi^2 analysis. If the authors want to keep it, they
need to argue why it is a more meaningful metric than RUWE, and
attempt to account for color and other parameters that impact
chi^2.

RESPONSE
----------
> We've done as suggested, and replaced this sub-plot with RUWE
> instead of reduced chi^2. The text in Section 4.3 and captions
> have been updated accordingly, and chi^2 is now mostly ignored.

******************************************
REFEREE COMMENT
----------

Section 4.4: Need a period between velocities and Further.

RESPONSE
----------
> Corrected, thank you.

******************************************
REFEREE COMMENT
----------

Section 5.2: That clear offset at (Pl-Ps)/Ps doesn't seem like a
coincidence, and the paper reads like the authors agree, but
couldn't explain it. I'm not sure I can help much here, although
it is interesting. Could this be that both stars are dippers and
one turns off while the other turns on by coincidence?  Spots and
hot-spot clusters can shift in phase as the magnetic fields
realign, which creates a new spot cluster at the same period but
a different phase. If it is circumstellar material, a tight
binary could be shepherding the material, but I think the systems
would have to be very tight for that, and it could not last for
such a long time.

I think this can be assessed statistically; given the scatter in
those points, the odds of lining up to any given line is small.
But without a physical explanation this is hard to test. Probably
this requires looking at many dippers as an ensemble. It is not
required for the primary analysis of this paper, so I will defer
to the authors if they want to dig deeper on this.

RESPONSE
----------
> Indeed, we are not sure whether or not it is a coincidence, but
> we thought it was interesting enough to mention, even though we
> do not have an interpretation.
>
> It seems unlikely in our view that both stars are showing dips.
> Our interpretation of the two smooth photometric signals is
> that one star is rotating at 11.98 hr, and the other star is
> rotating at 10.76 hr.
>
> All of the dips occur with a periodicity of 10.76 hr. (If we
> make Figure 6 but instead phasing by P_long, it yields a mess
> of points).  Therefore all the dips are presumably associated
> with the star that has that rotation period.

******************************************
REFEREE COMMENT
----------

Section 5.3.2: The difference between the fraction of transient
flux dip stars and the binary fraction is more significant than
it appears to the authors.  High mass ratio binaries would not be
included, as they are too faint to detect the second period.
Further, at the distance of Upper Sco (~150pc) and assuming
Stauffer et al. 2017 would notice/separate two stars wider than a
K2 pixel (4"), the presence of two periods would not include
binaries wider than 600pc.  Eyeballing the Winters et al. 2019
distribution suggests ~20% of binaries are beyond this. Stauffer
et al. 2017 also probably checked against existing imaging data
for obvious binaries down to 1-2", so likely even more are
getting cut.

RESPONSE
----------
> We have updated the text to include this point -- thank you for
> making it.
> 
> We have also updated the text to reflect our hesitation about
> whether or not the phenomenon is associated with binarity,
> primarily because multiple photometric periods do not
> necessarily imply binarity.

******************************************
REFEREE COMMENT
----------

Section 5.4: The authors conclude this subsection with a
reasonable argument that this isn't a protoplanet. Does this same
argument not rule out gas/debris at the Keplerian coronation
radius? What could maintain such clouds for years/decades and
then have them jump phase but at the same period? It's hard to do
this if it is not somehow physically attached to the star in some
what (e.g., spots can do this).

RESPONSE
----------
> We have modified the last paragraphs of Section 5.4 to reflect
> an updated preference that we came to after submission, for the
> clumpy torus model over the spot occultation model. The reasons
> for the preference are given in the text: 1) the spot
> occultation model seems more geometrically contrived than we
> initially thought, and 2) the Halpha excess seen by
> Johns-Krull+16 (and also in unpublished observations by
> Stauffer of scallop shells) seems likely to require hydrogen in
> orbit near the corotation radius.
>
> In response to the question of: "why can the shifting phase be
> explained more easily by dust/gas clumps than the
> protoplanet?", we now refer D'Angelo and Spruit 2012 (see in
> particular their Section 1), which discusses the possibility of
> cyclic accretion, with periods of "stalled" mass accumulation
> outside the corotation radius alternating with accreting
> phases.

******************************************
ADDITIONAL CHANGES

A few additional changes are noted in the "List of Changes"
appended at the end of the manuscript.

We modified the text throughout to clarify that Stauffer+ 2017
favored the explanation of a clumpy torus of gas and dust at the
Keplerian co-rotation radius, rather than our initial
(incomplete) description of "clouds of gas" at the same location.

We also modified the labels of the axes in Figure 5a for clarity.

We have also added a few citations that we omitted in the
original manuscript.